GEE models

### 2017-10-14 doc/esm.Rmd
Added more explanation of GEE to manuscript + esm

Generalized estimating equations (GEE) is a technique similar to mixed effects 
modelling, except it calculates the marginal population estimates compared to the
subject-specific estimates in mixed effects models. GEE is well suited to and
commonly used in longitudinal cohort studies, especially given its capacity to
handle missed visits. GEE is able to handle longitudinal data due to the
estimating of a working correlation matrix, which was chosen based on
quasi-likelihood information criteria (QIC). The auto-regressive of order 1
(AR1) matrix was chosen for the GEE models as it had the best model fit accessed
using QIC, though other matrices (eg. exchangeable) had similar fit (data not
shown). One main advantage of using methods such as GEE is the inclusion of the
working correlation matrix that takes into account the inherent covariance
between values measured at different time points when computing model estimates
and standard errors. AR1 is the most appropriate correlation matrix for
longitudinal data since the correlation matrix assumes lower intercorrelation of
outcomes within an individual as time passes. For example, BMI measured at the
baseline visit would be more highly correlated to BMI measured at the 3rd year
visit than compared to the 6th year visit.

>>>>

Generalized estimating equations (GEE) is a technique similar to mixed effects 
modelling, except it calculates the marginal population estimates compared to the
subject-specific estimates in mixed effects models. GEE is well suited to and
commonly used in longitudinal cohort studies, especially given its capacity to
handle missed visits and for the correlation of variables across time. 
GEE is able to handle the correlation of longitudinal data due to the
estimating of a working correlation matrix, which was chosen based on
quasi-likelihood information criteria (QIC; more detail on QIC below). The auto-regressive of order 1
(AR1) matrix was chosen for the GEE models as it had the best model fit accessed
using QIC, though other matrices (eg. exchangeable) had similar fit (data not
shown). While the choice of the correlation matrix is important, it should be noted
that even with misspecification GEE still computes robust estimates and standard errors.
One main advantage of using methods such as GEE is the inclusion of the
working correlation matrix that takes into account the inherent covariance
between values measured at different time points when computing model estimates
and standard errors. AR1 is the most appropriate correlation matrix for
longitudinal data since the correlation matrix assumes lower intercorrelation of
values within an individual as time passes. For example, BMI measured at the
baseline visit would be more highly correlated to BMI measured at the 3rd year
visit than compared to the 6th year visit.

### 2017-10-14 doc/manuscript.Rmd
Added more explanation of GEE to manuscript + esm

For the primary analysis, generalized estimating equation (GEE) models 
[@Zeger1986a] were used to determine the longitudinal associations between the 
outcome variables and the predictor variables. Given the longitudinal design an
auto-regressive of order 1 (AR1) correlation matrix was chosen for the GEE
models, though other matrices (eg. exchangeable) had similar fit (data not
shown). GEE is well suited to longitudinal cohort studies given it's capacity to
handle missed visits. The predictor variables and continuous covariates were
scaled (mean centered and standardised). The NEFA variables were classified as 
*time-independent* (held constant) as they were measured only at the baseline 
visit, while the outcome variables were set as *time-dependent*. No imputation
was conducted on missing values and incident diabetes cases were excluded from 
the analysis.

Covariate selection was based on previous literature, directed acyclic graph 
[@Greenland1999a;@Textor2011a] recommendations (see `r cite_sf("dag_isi")` and 
`r cite_sf("dag_issi2")`), and quasi-likelihood information criterion (QIC). See
the ESM Methods for more detail on QIC and the directed acyclic graphs. 
`r cite_st('qic')` shows the covariates compared using QIC. We selected the
model (M9) that had the fewest covariates and that had similar fit between the
outcome measures. The final GEE model included the covariates years from baseline
(time), sex (female or male), ethnicity (European ancestry or non-European 
ancestry), baseline age, WC, ALT, and MET (physical activity). Sex, ethnicity, 
and baseline age were time-*independent* variables. See the ESM Methods for more
detail on the covariate inclusion. After scaling,
and log-transforming the model variables, then exponentiating the GEE estimates,
the interpretation of the GEE results is as an expected percent difference in 
the outcome variable for every standard deviation (SD) increase in the predictor
variable given the covariates are held constant. With the main effect 
associations (not with a time interaction), this interpretation is applied 
consistently at each time point and thus represents associations of baseline 
NEFA for each future value of the outcomes. Since TAG is a risk factor for
diabetes and since NEFA contribute to TAG production [@Barrows2006a], TAG may
act as a mediator between NEFA and the outcomes. To determine the role of TAG in
the association between NEFA and the outcomes, TAG was included in the GEE model
in a sensitivity analysis. Lastly for the GEE models, we tested for an 
interaction with sex, ethnicity, or time by the predictor term for each outcome 
variable.

>>>>

For the primary analysis, generalized estimating equation (GEE) models 
[@Zeger1986a] were used to determine the longitudinal associations between the 
outcome variables and the predictor variables. Given the longitudinal design an
auto-regressive of order 1 (AR1) correlation matrix was chosen for the GEE
models, though other matrices (eg. exchangeable) had similar fit (data not
shown). GEE is well suited to longitudinal cohort studies given it's capacity to
handle any missed visits and for the correlation of variables between time points. 
See the ESM Methods for more detail on GEE.
The predictor variables and continuous covariates were
scaled (mean centered and standardised). The NEFA variables were classified as 
*time-independent* (held constant) as they were measured only at the baseline 
visit, while the outcome variables were set as *time-dependent*. No imputation
was conducted on missing values and incident diabetes cases were excluded from 
the analysis.

Covariate selection was based on previous literature, directed acyclic graph 
[@Greenland1999a;@Textor2011a] recommendations (see `r cite_sf("dag_isi")` and 
`r cite_sf("dag_issi2")`), and quasi-likelihood information criterion (QIC) [@Pan2001a]. See
the ESM Methods for more detail on QIC and the directed acyclic graphs. 
`r cite_st('qic')` shows the covariates compared using QIC. We selected the
model (M9) that had the fewest covariates and that had similar fit between the
outcome measures. The final GEE model included the covariates years from baseline
(time), sex (female or male), ethnicity (European ancestry or non-European 
ancestry), baseline age, WC, ALT, and MET (physical activity). Sex, ethnicity, 
and baseline age were time-*independent* variables. See the ESM Methods for more
detail on the covariate inclusion. After scaling,
and log-transforming the model variables, then exponentiating the GEE estimates,
the interpretation of the GEE results is as an expected percent difference in 
the outcome variable for every standard deviation (SD) increase in the predictor
variable given the covariates are held constant. With the main effect 
associations (not with a time interaction), this interpretation is applied 
consistently at each time point and thus represents associations of baseline 
NEFA for each future value of the outcomes. Since TAG is a risk factor for
diabetes and since NEFA contribute to TAG production [@Barrows2006a], TAG may
act as a mediator between NEFA and the outcomes. To determine the role of TAG in
the association between NEFA and the outcomes, TAG was included in the GEE model
in a sensitivity analysis. Lastly for the GEE models, we tested for an 
interaction with sex, ethnicity, or time by the predictor term for each outcome 
variable.
